---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
data <- read.csv(unzip("activity.zip"))
  

```



## What is mean total number of steps taken per day?
```{r}
daily_steps <- tapply(data$steps,data$date,sum,na.rm=TRUE)
```

Histogram of daily steps:
```{r}
hist(daily_steps,n=20)
```

Mean daily steps:
```{r}
mean(daily_steps)
```
Median daily steps:
```{r}
median(daily_steps)
```


## What is the average daily activity pattern?
Average each interval
```{r}
daily_activity <- tapply(data$steps,data$interval,mean,na.rm=T)
```


Plot histogram
```{r}
plot(daily_activity,type="l",ylab = "Steps",xlab = "Interval")
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
names(daily_activity)[ which(daily_activity ==  max(daily_activity)) ]
```


## Imputing missing values
How many rows have at least one missing Value?

```{r}
sum(!complete.cases(data))
```

All missing values are in the steps column, as can be seen from
```{r}
summary(data)
```

Replace missing values by sampling randomly from the same intervals with non missing values

```{r}
data2 <- data
set.seed(20)

for (i in seq(nrow(data))){
  if (is.na(data2[i,1])){
    data2[i,1] <- sample(data[data$interval==data[i,3] & !is.na(data$steps),1],1)
  }
}
```

How did filling the NAs change our statistics?

```{r}
daily_steps <- tapply(data2$steps,data2$date,sum,na.rm=T)
```

Histogram of daily steps:
```{r}
hist(daily_steps,n=20)
```

Mean daily steps:
```{r}
mean(daily_steps)
```
Median daily steps:
```{r}
median(daily_steps)
```

These values increased.
Days with missing values where earlier interpreted as having zero steps. This reduces mean and median.


## Are there differences in activity patterns between weekdays and weekends?

```{r}
library(lubridate)
data2$wdays <- as.factor(sapply(wday(ymd(as.character(data2$date))),function(x) if (x==1 | x==7){"weekend"}else{"weekday"}))
daily_activity <- data.frame(tapply(data2$steps,list(data2$interval,data2$wday),sum,na.rm=T))

daily_activity$interval <- rownames(daily_activity)

library(tidyr)

daily_activity <- gather(daily_activity,key="wdays",value = "Steps",-interval)


library(ggplot2)

daily_activity$interval <- as.numeric(daily_activity$interval)

qplot(interval,Steps,data=daily_activity,facets = wdays~.,geom ="line")


```



